---
#title: "Macroenometrics (44343)"
output:
  html_document:
    theme: united
    highlight: monochrome 
    self_contained: yes
---


```{r chunk_setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE, 
                      cache = FALSE, cache.path = "/caches/", comment = "#>",
                      #fig.width = 7, fig.height= 7,   
                      #out.width = 7, out.height = 7,
                      collapse = TRUE,  fig.show = "hold",
                      fig.asp = 7/9, out.width = "75%", fig.align = "center")
```

```{r options_setup, echo = FALSE}
options(scipen = 999) #- para quitar la notacion cientifica
```

<br><br><br>

### VAR models (4th topic of Macroenometrics (44343))

-------------------

<br>


Hi students! Wellcome to the course website!!

<br>

This is the website for the 4th topic of the 2019-2020 Macroeconometrics course (Master in Economics) at the University of Valencia. Mariam has delivered already the first 3 topics and I’m in charge of the last and 4th: **VAR models**. The repo to build this web is hosted [here](https://github.com/perezp44/VAR-models-web/){target="_blank"} on Gitbub.

**Topic summary:** The topic is an introduction to structural VAR models, one of the principal research toolkit in macroeconomic research.

As you know the classes were face-to-face but then the covid-19 ... I hope all this will pass soon!! We will try our best lo learn as much as possible about VAR models!!


For more information about the mini-course, please visit the [`Course "plan"`](https://perezp44.github.io/VAR-models-web/01-Course-plan.html) section of the website.


<br>


-- Pedro J. Pérez


<br><br><br><br>
  