---
title: "My VAR model: SPAIN"
author: "Your name"
date: "`r format(Sys.time(), '%d %B, %Y')`" 
output: 
  html_document:
    theme: simplex
    highlight: pygments 
    number_sections: no
    toc: true
    toc_float: true
    #code_folding: show
    self_contained: TRUE
    code_download: true
---


```{r, include = FALSE}
#- definiendo opciones globales para los chunks de código R
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE, 
                      cache = FALSE, cache.path = "/caches/", comment = "#>",
                      #fig.width = 7, fig.height= 7,   
                      #out.width = 7, out.height = 7,
                      collapse = TRUE,  fig.show = "hold",
                      fig.asp = 7/9, out.width = "60%", fig.align = "center")
```


```{r, echo = FALSE}
options(scipen = 999) #- para quitar la notacion cientifica
```


```{r, echo = FALSE}
#- loading the packages we are going to use
library(tidyverse)     #- install.packages("tidyverse")
library(dygraphs)      #- install.packages("dygraphs")
library(DT)            #- install.packages("DT")
#PACKAGES <- c("devtools","ggplot2", "tidyr", "dplyr","knitr", "curl", "vars", "forecast", "dygraphs", "eurostat", "rmdformats", "DiagrammeR", "rio", "tidyverse")  #- packages to load

```


<br><br>

# 1. Intro

As the 20% of my mark for Macroeconometrics depends on the VAR assignment and my participation in the e-meetings, I'm going to start doing my assignment. 

For my assignment, I'm going to estimate a VAR model with Eurostat data. I downloaded the data using the `to-download-my_data.Rmd` file   the `xx/04/2020`. Big thanks to the the `eurostat` package.

And remember, as the lecturer said repeatedly during the e-meetings, that there are some rules to deliver the assignment to him:

- you have to deliver the assignment by mail to `pedro.j.perez@uv.es` **BEFORE the 15/05/2020**.   
 
- you should attach in you mail a **single .zip file**.   

- this single .zip file has to contain your **assignment as an Rproject**.   

- Remember that you have to name the folder of you Rproject like: <VAR_country_1st-surname_name>. That is, if for instance, in your assignment you have used data from `Spain` and your name would be "Pedro Pérez Vázquez", then, the folder of your Rproject should be named: `VAR_perez_pedro_spain`. Please, all in lower case, except VAR, and **no spaces, no punctuation marks** in the name of the folder.   

- Then, if the name of the folder of your Rproject is `VAR_perez_pedro_spain` then the .zip file you have to send me by e-mail should be named: `VAR_perez_pedro_spain.zip`.   

<br>

---------------------------------------

<br>

# 2. Data preparation

As we are going to use the `vars` package to estimate and analyse a structural VAR model, I need to transform my data to a format suitable for the `vars` package. 

<br>

## Loading my data

Obviously, first I have to load my data. As I have saved the file with my data in a the folder called `datos` inside the main folder of the Rproject, THEN, to load my data I have to do:


```{r, eval = FALSE}
#- remenber that you should change the name of the path to the file
df <- read_csv("./datos/my_data_ES.csv") 
```

Other way to load the data. This second way to load the data is preferable. 


```{r}
df <- read_csv(here::here("datos", "my_data_ES.csv")) 
```

We are going to use only 3/4 the variables: Vol, Vol_15, Vol_us & time


```{r}
data <- df %>% dplyr::select(time, Vol_us, Vol_15, Vol)
```






<br>

## Creating the `ts` objects

As the `vars` package need that the data are in a special format (`ts`), then, I have to convert my data to the `ts` format

To do that, first we have to look at the common sample of our 3 series:

Dropping rows with NAs to obtain the The common sample for my data.

```{r}
data <- data %>% drop_na()
```

calculating the start and end of the common sample of my data:
  
```{r}
data_tt <- data %>% mutate(time = as.character(time)) %>% tidyr::separate(time, c("year", "quarter", "day"),"-")
data_tt <- data_tt %>% mutate(year = as.numeric(year)) 
data_tt <- data_tt %>% mutate(quarter= case_when(
                                        quarter  == "01"  ~ 1,
                                        quarter  == "04"  ~ 2,
                                        quarter  == "07"  ~ 3,
                                        quarter  == "10"  ~ 4))
start_year    <- data_tt$year[1] 
start_quarter <- data_tt$quarter[1] 
end_year      <- data_tt$year[nrow(data_tt)] 
end_quarter   <- data_tt$quarter[nrow(data_tt)] 
```


OK, now we can create the `ts` object needed to work with the `vars` package.

```{r}
#- Creating time series: `ts` object
ser <- list(data$Vol, data$Vol_15, data$Vol_us)
series <- lapply(ser, ts, start = c(start_year ,start_quarter) , end = c(end_year, end_quarter) , frequency = 4)
names(series) <- c("Vol","Vol_15","Vol_us")
```

<br>

Creating logs and first log-differences of the time series

```{r}
series.l  <- lapply(series, log)
series.dl <- lapply(series.l, diff)
```

<br>

Selecting, binding and naming the 3 series we are going to use:

```{r}
PIBs <- cbind(series[[1]], series[[2]], series[[3]])
colnames(PIBs) <- c("Vol","Vol_15","Vol_us")
```

<br>

A dynamic graph (with the `dygraph` package):

```{r}
dygraph(PIBs, main = "Evolution of GDP levels")  %>%  dyRangeSelector() %>%
dyOptions(colors = RColorBrewer::brewer.pal(3, "Set2")) 
```

<br><br><br>

Making a graph for the series in log-first-differences: (1-L)logGDP

```{r}
d.PIBs <- cbind(series.dl[[1]], series.dl[[2]], series.dl[[3]])
colnames(d.PIBs) <- c("Vol","Vol_15","Vol_us")
dygraph(d.PIBs, main = "Evolution of (1-L) log GDP")  %>%  dyRangeSelector() 
```


<br>

-------------------------------------------------------

<br><br>


 
# 3. Estimation of the VAR model

OK, I have to estimate the VAR, let's load the vars package .....



<br><br><br><br>
